---
title: "Text analysis"
output: html_document
date: "2024-03-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(quanteda)
library(quanteda.textstats)
library(quanteda.textplots)
library(quanteda.sentiment)
library(stringr)
library(stm)
library(lubridate)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(reshape2)
library(patchwork)
```

## Data Loading and Preprocessing
Load the data, which include the combined blog articles of the four organisations.
```{r}
articles <- read.csv("article_text.csv")
articles$Organisation <- as.factor(articles$Organisation)

# Creating a day of the year column
articles$Date <- dmy(articles$Date)
articles$day_of_year <- yday(articles$Date)

summary(articles)
head(articles)
```

Preprocess the text data.
```{r}
# Original articles text
texts <- articles$Text

# Removing lengthy disclaimers from Rainbow Migration directly from the text
disclaimer_pattern <- "Disclaimer The content of this website is provided for general information only and should not be relied on as legal advice. Rainbow Migration disclaims any liability resulting from reliance on such information. You are strongly advised to seek professional legal advice from a qualified immigration solicitor.  hello@rainbowmigration.org.uk020 7922 78117-14 Great Dover St,London SE1 4YR © 2024 Rainbow Migration. All rights reserved"

# Use str_remove_all to remove patterns from each text
cleaned_texts <- str_remove_all(texts, disclaimer_pattern)

# Create corpus from cleaned texts, set 3 docvariables associated with each article/document
corp <- corpus(cleaned_texts,
               docvars = data.frame(organisation = articles$Organisation,
                                    title = articles$Title,
                                    date = articles$Date,
                                    day_of_year = articles$day_of_year))

# create the dfm
dfm_ungrouped <- corp %>% 
  tokens(remove_punct = TRUE, remove_numbers = TRUE, remove_symbols = TRUE,
         remove_url = TRUE) %>%
  tokens_remove(stopwords("en")) %>%
  tokens_wordstem() %>%
  tokens_ngrams(n = 1:2) %>%
  dfm() %>%
  dfm_trim(min_termfreq = 5)

# group the dfm for descriptive analysis
dfm_grouped <- dfm_group(dfm_ungrouped, groups = articles$Organisation)

# Inspect the ungrouped and grouped dfm
topfeatures(dfm_grouped, n = 50)
dfm_grouped
topfeatures(dfm_ungrouped, n = 50)
dfm_ungrouped
```

## Descriptive analysis
Create a grouped wordcloud to visualise the most frequent words used by each organisation, so to have a brief picture of the frequently used words by each organisation in their blog articles.
```{r, warning=FALSE}
textplot_wordcloud((dfm_grouped), 
                   comparison = TRUE, 
                   groups = articles$Organisation,
                   min_count = 10,
                   max_words = 320,
                   color = c("red", "blue", "purple", "green"))
```

Structural topic modelling (STM) is a method that combines topic modelling and regression analysis to identify topics and their relationship with metadata. In this case, we will use STM to identify the topics in the blog articles and their relationship with the organisation that published them.
```{r}
# convert the quanteda object into a stm input
stm_input <- convert(dfm_ungrouped, to = "stm")
# quantitative diagnostics
k_search_output <- searchK(stm_input$documents, stm_input$vocab,
                           K = c(5, 6, 7, 8, 9), data = stm_input$meta,
                           verbose = FALSE, heldout.seed = 123)
plot(k_search_output)
k_search_output
```

K = 8 seems to be the sweet spot with the highest held-out likelihood, suggesting good model fit.
K = 8 also has improved semantic coherence compared to K = 5, 6, and 7, indicating that the topics are meaningful.
Although K = 9 has slightly better residuals and a higher lower bound, the semantic coherence drops, which might indicate that the topics at K = 9 are less interpretable.

Given these diagnostics, K = 8 appears to be the most suitable number of topics because it provides the best balance across the various measures. It has good predictive accuracy (held-out likelihood), a good balance of topic interpretability (semantic coherence), and a reasonable model fit (residuals and lower bound), without the drop in coherence observed at K = 9.

```{r}
stmodel <- stm(documents = stm_input$documents, vocab = stm_input$vocab,
                     K = 8, prevalence =~ organisation + s(day_of_year),
               data = stm_input$meta, verbose = FALSE, init.type = "LDA", seed = 123) # initialise the structural topic model with LDA as it works better in this scenario
plot(stmodel)
```
```{r}
# plot wordclouds for a detailed examination of each topic, so to label topics

cloud(stmodel, topic = 1, scale = c(2, .25)) # Refugee Safety and Essential Support
cloud(stmodel, topic = 2, scale = c(2, .25)) # Organisation's Progress and Opportunities
cloud(stmodel, topic = 3, scale = c(2, .25)) # Organisational Leadership for Migrant Support
cloud(stmodel, topic = 4, scale = c(2, .25)) # Diversity and Historical Acknowledgment
cloud(stmodel, topic = 5, scale = c(2, .25)) # Migrant Data and Legal Status
cloud(stmodel, topic = 6, scale = c(2, .25)) # Migrants' Right-to-Work
cloud(stmodel, topic = 7, scale = c(2, .25)) # Women's Experiences in Asylum Seeking
cloud(stmodel, topic = 8, scale = c(2, .25)) # LGBTQI+ Experiences in Asylum Seeking and Detention

```
Plot to see which topics are most prevalent in each organisation's blog articles.
```{r}
# estimate the effect of organisation and day_of_year on topic prevalence
effect_estimates <- estimateEffect(1:8 ~ organisation + s(day_of_year), stmodel, meta = stm_input$meta)

# Get document-topic probabilities directly from the model
topic_probabilities <- as.data.frame(stmodel$theta) # theta matrix contains the document-topic probabilities

# Prepare the data by adding document identifiers and melting to long format
topic_probabilities$document <- rownames(topic_probabilities)
doc_topic_org <- tidyr::pivot_longer(topic_probabilities, -document, names_to = "topic", values_to = "prob")

# labels for the topics
topic_labels <- c("Refugee Safety and Essential Support", "Organisation's Progress and Opportunities", "Organisational Leadership for Migrant Support", "Diversity and Historical Acknowledgment", "Migrant Data and Legal Status", "Migrants' Right-to-Work", "Women's Experiences in Asylum Seeking", "LGBTQI+ Experiences in Asylum Seeking and Detention")

# Convert the topic number to a factor with levels corresponding to the numbers and labels as the topic labels
doc_topic_org$topic <- factor(doc_topic_org$topic, levels = paste("V", 1:8, sep=""), labels = topic_labels)

# Add a document identifier to the metadata
stm_input$meta$document <- seq_len(nrow(stm_input$meta))

# Merge with metadata to include organisation information
doc_topic_org <- merge(doc_topic_org, stm_input$meta, by = "document")

# Corrected aggregation step
topic_prevalence_by_org <- aggregate(prob ~ topic + organisation, data = doc_topic_org, mean)

# Plot aggregated topic prevalences for each organisation
ggplot(topic_prevalence_by_org, aes(x = reorder(topic, -prob), y = prob, fill = organisation)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(x = "Topic", y = "Average Topic Prevalence", title = "Topic Prevalence by Organisation") +
  coord_flip() +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))
```

Plot the discussion of topics over the specified timeframe.
```{r}
# Set up the plotting area to have 2 rows and 4 columns
par(mfrow=c(2,4))

# Number of topics
n_topics <- length(topic_labels)

# Initialise a list to store plots
plot_list <- vector("list", n_topics)

# Create a plot for each topic
for (i in 1:n_topics) {
  plot_list[[i]] <- plot(effect_estimates, "day_of_year", method = "continuous", topics = i,
                         model = stmodel, printlegend = FALSE, xaxt = "n", xlab = "2022-2023",
                         main = topic_labels[i])
  
  # Assuming you have a sequence of dates for the x-axis (as in your example)
  monthseq <- seq(from = as.Date("2022-12-01"), to = as.Date("2023-12-31"), by = "month")
  monthnames <- months(monthseq)
  axis(1, at = as.numeric(monthseq) - min(as.numeric(monthseq)), labels = monthnames, las=2) # `las=2` makes labels perpendicular
}

```

## Sentiment Analysis
```{r}
data(data_dictionary_NRC)
# weight the grouped dfm in proportion
dfm_grouped_weighted <- dfm_weight(dfm_grouped, scheme = "prop")
# apply the NRC dictionary to the weighted dfm
emotions <- dfm_lookup(dfm_grouped_weighted, dictionary = data_dictionary_NRC)
emotions

# Assuming emotions is your dfm after applying the NRC dictionary
# First, create a subset for negative and positive sentiments
neg_pos <- emotions[, c("negative", "positive")]

# Normalize so that each row sums to 1
neg_pos <- neg_pos / rowSums(neg_pos)

# Do the same for the other eight emotions
other_emotions <- emotions[, c("anger", "anticipation", "disgust", "fear", "joy", "sadness", "surprise", "trust")]

# Normalize these as well
other_emotions <- other_emotions / rowSums(other_emotions)

# Combine the two normalized sets back into one dfm
normalised_emotions <- cbind(neg_pos, other_emotions)
normalised_emotions
```
```{r}
# Convert DFM to matrix
emotions_matrix <- as.matrix(normalised_emotions)

# Melt the matrix into a long data frame
emotions_long <- melt(emotions_matrix, value.name = "sentiment_score")

# Correctly match the organization names to the emotions_long data frame
# Get the number of rows for each document (number of features per doc)
rows_per_doc <- nfeat(emotions)

# Repeat the organization names according to the number of features per document
emotions_long$organisation <- rep(docvars(emotions, "organisation"), times = rows_per_doc)

# Since we're using melt, which doesn't assign column names for document and feature, let's add them
colnames(emotions_long)[1:2] <- c("document", "feature")

# Filter for only negative and positive emotions
emotions_neg_pos <- emotions_long %>%
  filter(feature %in% c("negative", "positive"))

# Plot negative vs positive sentiments by organisation
ggplot(emotions_neg_pos, aes(x = organisation, y = sentiment_score, fill = feature)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x = "Organisation", y = "Sentiment Score", title = "Negative vs Positive Sentiment by Organisation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
# Filter out negative and positive emotions to focus on the other emotions
emotions_others <- emotions_long %>%
  filter(!feature %in% c("negative", "positive"))

# Plot all other emotions by organisation
ggplot(emotions_others, aes(x = organisation, y = sentiment_score, fill = feature)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Organisation", y = "Sentiment Score", title = "Other Emotions by Organisation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "bottom")  # Adjust legend position for readability
```

```{r}
# Weight the ungrouped DFM by count
dfm_ungrouped_weighted <- dfm_weight(dfm_ungrouped, scheme = "count")

# Convert DFM to a matrix, then to a data frame
emotion_matrix <- as.matrix(text_emotions)
emotion_df <- as.data.frame(emotion_matrix)

emotion_df

# Ensure document identifiers are properly set as row names in the DFM
row.names(emotion_df) <- docnames(text_emotions)

# Convert row names to a column for document identifiers
emotion_df$document <- row.names(emotion_df)

# Now, pivot only the sentiment score columns to long format
emotion_long <- pivot_longer(emotion_df, 
                             cols = -document, 
                             names_to = "emotion", 
                             values_to = "sentiment_score")

# Merge with metadata
doc_metadata <- data.frame(
  document = docnames(dfm_ungrouped),
  organisation = docvars(dfm_ungrouped, "organisation"),
  date = as.Date(docvars(dfm_ungrouped, "date"), format = "%Y-%m-%d")
)

emotion_long <- left_join(emotion_long, doc_metadata, by = "document")

# Ensure date is correctly formatted
# emotion_long$date <- as.Date(emotion_long$date)

# Specify the exact organization names
organisations <- c("Freedom from Torture", "Migrants' Rights Network", 
                   "Rainbow Migration", "Women for Refugee Women")

# Initialize an empty list for storing plots
plot_list <- list()

# Iterate over organizations to create plots
for (org in organisations) {
  # Filter data for the current organization
  filtered_data <- emotion_long %>%
    filter(organisation == org) %>%
    filter(emotion %in% c("joy", "trust", "fear", "surprise", "anger", "anticipation", "disgust", "sadness"))

  # Create plot for current organization
  p <- ggplot(filtered_data, aes(x = date, y = sentiment_score, color = emotion)) +
    geom_line() +
    labs(title = paste("Sentiment Over Time for", org),
         x = "Date", y = "Sentiment Score") +
    scale_color_viridis_d() +
    theme_minimal() +
    theme(legend.position = "bottom", legend.title = element_blank()) +
    scale_x_date(date_labels = "%b %Y", date_breaks = "1 month")

  # Add the plot to the list
  plot_list[[org]] <- p
}

# Combine the plots using patchwork
combined_plot <- wrap_plots(plot_list, ncol = 2)

# Print the combined plot
print(combined_plot)

```